Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-03-04 draft.tex
.

Effenberger et al. identified the presence of inhibitory STDP as a necessary mechanism for the development of this feature, in contrast to our model. On a deeper explanatory level, the authors argue that inhibitory STDP is indirectly involved: Though it is known that inhibitory STDP can balance network activity \cite{Vogels_2011}, for some excitatory neurons in the network, presynaptic inhibitory neurons can not fully compensate the above-average firing rate of the postsynaptic neuron. Furthermore, synapses from highly active excitatory presynaptic cells to postsynaptic excitatory neurons with lower activity are known to be subject to long-term potentiation \cite{Sjoestroem_Syn_Plasticity_2001,Feldman_STDP_2012}. These two effects combined then give rise to the strong mean outgoing weights of this subgroup of neurons. With these causal relations in mind, we can argue that diffusive homeostasis effectively embodies a similar functional role as inhibitory STDP in \cite{Effenberger_2015} by allowing for the presence of excitatory cells with above-average activity. We tested this relationship by plotting the mean of excitatory outgoing weights, sum of outgoing weights and the out degree against the average firing rate, see Figure \ref{Out_Weight_vs_F}. A strong heterogeneity of firing rates allows for the development of few neurons with comparably strong mean outgoing weights, while the distribution of mean weights resulting from a narrow distribution is limited to a smaller range. Interestingly, the heterogeneity of out weights is most prominent when looking at the total sum. While this should lead to the conclusion that this is the case because an increased firing rate increases both, mean and number of outgoing weights, the latter can not be strictly verified based on Figure \ref{Out_Weight_vs_F} (D). In consequence, we also did not find a strong direct correlation between mean outgoing weights and out degree, as shown in Figure \ref{Mean_Weight_vs_Out_Degree}. Still though, it is interesting to note that diffusive homeostasis, in contrast to the other shown variants, diffusive homeostasis at least has a correlation coefficient of $\rho_{\rm Diff.} =  0.4$, especially in contrast to $\rho_{\rm Inst. Diff.} =  -0.23$, though instant diffusion leads to similarly spread firing rates.

@@

\newpage
\section{Further Analysis of the Firing Rate Distribution}
The results of section \ref{Fir_Dist_Section} were positive on a preliminary level regarding our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}\label{Mean_Field_Review_Section}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_{t,i}}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the mixture between single-neuron homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their overall mean.

The spiking recurrent network was replaced by a set of non-interacting non-spiking neurons whose individual mean firing rates were calculated by
\begin{align}
\phi_i(\mu_i,\sigma_i,V_{t,i}) &= \left[ \sqrt{\pi}\tau_m \int_{x_-}^{x_+} dx e^{x^2} \mathrm{erfc}(-x) \right]^{-1} \label{LIF_Fir_Rate_with_Noise1_copy}\\
x_- &= (V_r-E_l-\mu_i\tau_m)/\sigma_i \label{LIF_Fir_Rate_with_Noise2_copy}\\
x_+ &= (V_{t,i}-E_l-\mu_i\tau_m)/\sigma_i \label{LIF_Fir_Rate_with_Noise3_copy}
\end{align}
which is the same formula we used in \eqref{LIF_Fir_Rate_with_Noise1}. Additionally, in accordance to \cite{Roxin_Firing_Rate_Distribution}, mean and standard deviation of the inputs were calculated by Sweeney et al. as
\begin{align}
\nu &= \langle \phi \rangle = \frac{\sum \phi_i}{N} \label{sweeney_self_consist_rate1} \\
\mu_i &= J_iC_i\nu \tau \label{sweeney_self_consist_rate2} \\
\sigma_i &= J_i\sqrt{C_i\nu  \tau} \label{sweeney_self_consist_rate3}
\end{align}
where $\nu$ is the mean population firing rate, $\phi_i$ the individual firing rate of neuron $i$, $\mu_i$, $\sigma_i$ and $V_{t,i}$ its synaptic input mean and standard deviation and intrinsic firing threshold respectively and $J_i$, $C_i$ and $\tau$ the neuron's mean synaptic efficacy, number of incoming neurons and the membrane time constant. Synaptic efficacy and number of incoming neurons were drawn randomly to match the statistics of the actual network topology. Self consistency was achieved by iterating through equations \eqref{LIF_Fir_Rate_with_Noise1_copy} and \eqref{sweeney_self_consist_rate1}--\eqref{sweeney_self_consist_rate1} until the desired precision of convergence was reached.

The authors claimed that this model reproduces results of the full network, in particular that the steady-state firing rate distribution spreads out due to a larger diffusion constant (or a corresponding larger $\alpha$).

@@

Figure \ref{dynamics_rate_threshold_dyn_mean_field_sweeney} depicts the resulting dynamics of excitatory rates and thresholds for $\mathrm{\alpha=\lbrace 0.4,0.8,0.9,1.0\rbrace}$. Apparently, all rates approach the same target rate of $\mathrm{3\,Hz}$ and thresholds move towards the same configuration for either choice of $\mathrm{\alpha}$ except for the limiting case of $\mathrm{\alpha=1}$. A significant difference only exists within the transient dynamics leading to the steady state. Roughly speaking, a smaller value of $\mathrm{\alpha}$ leads to a faster relaxation.
\begin{figure}
%%% plot_sim_mean_field_sweeney.py
\includegraphics[width=\textwidth]{../../plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates (A) and thresholds (B) of excitatory population of 400 Neurons and different values of $\mathrm{\alpha}$ (see equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}).}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}
These findings led us to the conclusion that---apart from the special case $\mathrm{\alpha=1}$, which is the equivalent to instantaneous diffusion---a mean-field model describing the diffusive signal as a mixture of individual activity and all neurons is not suitable for explaining the existence of stable heterogeneous distributions of rates for a broad range of diffusion constants. We did not intent to debase the validity of results of the mean-field model of Sweeney et al. However, we could show that the steady-state solution of this simplified model will result in the same firing rate $\phi_0$ for all neurons, given any value of $\mathrm{\alpha}$ except $\mathrm{\alpha=1}$. This can be seen by setting the left hand side of Equation \eqref{diff_hom_simpl_sweeney} to $0$ (which is necessarily the case in the steady state) and rearranging the equation:
\begin{equation}
(\alpha-1)\frac{\phi_i - \phi_0}{\phi_i} = \alpha \frac{\langle \phi \rangle - \phi_0}{\langle \phi \rangle}
\label{diff_hom_simpl_sweeney_2}
\end{equation}
The right term of the equation is the same for all neurons. Since the left term is monotonically increasing as a function of $\phi_i$ as long as $\mathrm{\alpha<1}$, only one specific solution $\phi_i = \Phi$ for all $i$ exists that equals the given term on the right. This implies $\langle \phi \rangle = \Phi$. Thus,
\begin{equation}
(\alpha-1)\frac{\Phi - \phi_0}{\Phi} = \alpha \frac{\Phi - \phi_0}{\Phi}
\label{diff_hom_simpl_sweeney_3}
\end{equation}
which is only fulfilled for $\Phi = \phi_0$.

@@

Equation \eqref{simple_NO_dyn_with_diff_equil} can be rewritten as
\begin{equation}
\left( \nabla^2 + \left( i\sqrt{\frac{\lambda}{D}}\right)^2\right) NO = \sum_{i} \delta^2(\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i})\cdot \frac{- \gamma \cdot r_i}{D}
\label{simple_NO_dyn_with_diff_equil_helmholtz}
\end{equation}
which is a two-dimensional Helmholtz equation with a superposition of rescaled Dirac functions. The solution of $NO$ is thus composed of a superposition of shifted and scaled versions of the Green's function of the differential operator on the left hand side of the equation. For each delta function $\delta^2(\mathbf{x}-\mathbf{x}_i)$, the solution is
\begin{equation}
NO_i(\mathbf{x}) = \frac{r_i \gamma}{2 \pi D} K_0 \left(|\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i}|\sqrt{\frac{\lambda}{D}} \right) \equiv r_i \psi_{\rm point}(|\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i}|)
\label{solution_diff_equil_bessel}
\end{equation}
where $\mathrm{K_0}$ is the zeroth modified Bessel function of the second kind \cite{Helmholtz_Solution_2d}. This solution reveals a fundamental problem of modeling the sources of NO-production as point sources: the fact that $\mathrm{K_0(x)}$ diverges to infinity for $\mathrm{x\rightarrow 0}$. It is merely due to the finite density of the numeric grid used for the simulation of diffusion that allows for a finite target value of concentration. Note that this problem only occurs in the two- or three-dimensional version of the differential equation, whereas in one dimension, the fundamental solution can be expressed as an exponential function with respect to the distance to the origin, resulting in a well-defined finite value at $\mathrm{x=0}$.

@@

By simply calculating all matrix elements of $\mathrm{\hat{\psi}}$ by means of Equation \eqref{interact_matrix_elements}, one would neglect the finite boundaries of the system, which would cause neurons close to the edge to ``bleed" into empty space. This in turn would cause the solution of \eqref{NO_equil_cond_interact_matrix_operator} to contain an over-representation of high firing rates since close-to-the-edge neurons would need to compensate for their lack of neighbors. Thus, we had to account for the boundary conditions used in the network simulation. We simulated the network mostly with Neumann boundary conditions as well as periodic boundaries, as described in section \ref{methods}. Both types can be modeled by extending the neurons' population through spatially shifted and mirrored versions of the base population (see Figure \ref{Bound_Cond_Patches}):
\begin{itemize}
\item Periodic boundary conditions are induced by copies of the neurons' positions shifted by $L\cdot (n_x,n_y)$, $n_x,n_y \in \mathbb{Z}$.
\item Zero flux through the boundaries can be achieved by copied positions being shifted by $L\cdot (n_x,n_y)$, $n_x,n_y \in \mathbb{Z}$ \textit{and} mirrored in $x$($y$)-direction if $n_x$($n_y$) is odd.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{../../plots/Boundary_Cond_Sketch.png}
\end{center}
\caption{Sketch for patches of copied (and mirrored) positions for periodic (A) and Neumann (B) boundary conditions.}
\label{Bound_Cond_Patches}
\end{figure}
Therefore, the entries of the operator for periodic boundary conditions $\hat{\psi}_{\rm per.}$ could be calculated by
\begin{equation}
\psi_{ij,\rm per.} = \sum_{n_x,n_y \in \mathbb{Z}} \psi_{\rm approx.} (|\mathbf{x}_{\mathrm{ neur.},i}-\mathbf{x}_{\mathrm{ neur.},i} - L \cdot (n_x,n_y)|) \; .
\label{Psi_entries_periodic_bound}
\end{equation}
For the Neumann boundary condition, we found
\begin{align}
\psi_{ij,\rm neum.} &= \sum_{n_x,n_y \in \mathbb{Z}} \psi_{\rm approx.} \left(|\mathbf{x}_{\mathrm{ neur.},i}-M(\mathbf{x}_{\mathrm{neur.},i}) - L \cdot (n_x,n_y)| \right) \label{Psi_entries_neumann_bound} \\
M(\mathbf{x}) &\equiv  
 \begin{pmatrix}
  (-1)^{n_x} & 0\\
  0 & (-1)^{n_y}
 \end{pmatrix}
\mathbf{x} + L \cdot (mod(\left| n_x\right|,2),mod(\left| n_y\right|,2)) \; . \label{Mirror_Operator}
\end{align}
Note that the shift was applied to the second position. Theoretically, $n_x$,$n_y$ are to be iterated over all integers and in this case it is irrelevant whether to shift the first or the second position. We obviously had to limit the amount of elements of the sum for our calculations, such that copies that are shifted further away can be neglected. For the sake of performance the number of terms in the sum was adjusted to suffice the given diffusion constant.

@@

These observations naturally led us to the question of how the correlation between predicted measured firing rates behaves in between the aforementioned limits. Especially, we were interested in the range of the diffusion constant for which our model provides a good description of the full spiking network's activity. Figure \ref{Corr_Coeff_vs_D} depicts the Pearson correlation coefficient of $\mathrm{f_{matrix \\, predict}}$ and $\mathrm{f_{sim}}$ against the diffusion constant used in the simulation. A relatively high correlation was obtained for a wide range of diffusion constants. However, we could see a general decline of correlation for larger diffusion constants. This trend was in line with the aforementioned limit of instant diffusion, namely a complete decorrelation between prediction and measurement, as well as the good agreement for $\mathrm{D= 10\, \mu m^2 ms^{-1}}$ shown in Figure \ref{Fir_Rate_Dist_Sim_vs_Mat}.

>>>>

Effenberger et al. identified the presence of inhibitory STDP as a necessary mechanism for the development of this feature, in contrast to our model. On a deeper explanatory level, the authors argue that inhibitory STDP is indirectly involved: Though it is known that inhibitory STDP can balance network activity \cite{Vogels_2011}, for some excitatory neurons in the network, presynaptic inhibitory neurons can not fully compensate the above-average firing rate of the postsynaptic neuron. Furthermore, synapses from highly active excitatory presynaptic cells to postsynaptic excitatory neurons with lower activity are known to be subject to long-term potentiation \cite{Sjoestroem_Syn_Plasticity_2001,Feldman_STDP_2012}. These two effects combined then give rise to the strong mean outgoing weights of this subgroup of neurons. With these causal relations in mind, we can argue that diffusive homeostasis effectively embodies a similar functional role as inhibitory STDP in \cite{Effenberger_2015} by allowing for the presence of excitatory cells with above-average activity. We tested this relationship by plotting the mean of excitatory outgoing weights, sum of outgoing weights and the out degree against the average firing rate, see Figure \ref{Out_Weight_vs_F}. A strong heterogeneity of firing rates allows for the development of few neurons with comparably strong mean outgoing weights, while the distribution of mean weights resulting from a narrow distribution is limited to a smaller range. Interestingly, the heterogeneity of out weights is most prominent when looking at the total sum. While this should lead to the conclusion that this is the case because an increased firing rate increases both, mean and number of outgoing weights, the latter can not be strictly verified based on Figure \ref{Out_Weight_vs_F} (D). In consequence, we also did not find a strong direct correlation between mean outgoing weights and out degree, as shown in Figure \ref{Mean_Weight_vs_Out_Degree}. Still though, it is interesting to note that diffusive homeostasis, in contrast to the other shown variants, at least has a correlation coefficient of $\rho_{\rm Diff.} =  0.4$, especially in contrast to $\rho_{\rm Inst. Diff.} =  -0.23$, even though instant diffusion leads to similarly spread firing rates.

@@

\newpage
\section{Further Analysis of the Firing Rate Distribution}
The results of section \ref{Fir_Dist_Section} were positive on a preliminary level regarding our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}\label{Mean_Field_Review_Section}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_{t,i}}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the mixture between single-neuron homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their population mean.

The spiking recurrent network was replaced by a set of non-interacting non-spiking neurons whose individual firing rates were calculated by
\begin{align}
\phi_i(\mu_i,\sigma_i,V_{t,i}) &= \left[ \sqrt{\pi}\tau_m \int_{x_-}^{x_+} dx e^{x^2} \mathrm{erfc}(-x) \right]^{-1} \label{LIF_Fir_Rate_with_Noise1_copy}\\
x_- &= (V_r-E_l-\mu_i\tau_m)/\sigma_i \label{LIF_Fir_Rate_with_Noise2_copy}\\
x_+ &= (V_{t,i}-E_l-\mu_i\tau_m)/\sigma_i \label{LIF_Fir_Rate_with_Noise3_copy}
\end{align}
which is the same formula we used in \eqref{LIF_Fir_Rate_with_Noise1}. Additionally, in accordance to \cite{Roxin_Firing_Rate_Distribution}, mean and standard deviation of the inputs were calculated by Sweeney et al. as
\begin{align}
\nu &= \langle \phi \rangle = \frac{\sum \phi_i}{N} \label{sweeney_self_consist_rate1} \\
\mu_i &= J_iC_i\nu \tau \label{sweeney_self_consist_rate2} \\
\sigma_i &= J_i\sqrt{C_i\nu  \tau} \label{sweeney_self_consist_rate3}
\end{align}
where $\nu$ is the mean population firing rate, $\phi_i$ the firing rate of neuron $i$, $\mu_i$, $\sigma_i$ and $V_{t,i}$ its synaptic input mean and standard deviation over time and intrinsic firing threshold respectively and $J_i$, $C_i$ and $\tau$ the neuron's mean synaptic efficacy, number of incoming neurons and the membrane time constant. Synaptic efficacy and number of incoming neurons were drawn randomly to match the statistics of the actual network topology. Self consistency was achieved by iterating through Equations \eqref{LIF_Fir_Rate_with_Noise1_copy} and \eqref{sweeney_self_consist_rate1}--\eqref{sweeney_self_consist_rate3} until the desired precision of convergence was reached.

The authors claimed that this model reproduces results of the full network, in particular that the steady-state firing rate distribution spreads out due to a larger diffusion constant (or corresponding to a larger $\alpha$).

@@

Figure \ref{dynamics_rate_threshold_dyn_mean_field_sweeney} depicts the resulting dynamics of excitatory rates and thresholds for $\mathrm{\alpha=\lbrace 0.5,0.9,1.0\rbrace}$. We ran simulations with other values of $\rm \alpha$ and found that all rates approach the same target rate of $\mathrm{3\,Hz}$ and thresholds move towards the same configuration for either choice of $\mathrm{\alpha}$ except for the limiting case of $\mathrm{\alpha=1}$. A significant difference only exists within the transient dynamics leading to the steady state. Roughly speaking, a smaller value of $\mathrm{\alpha}$ leads to a faster relaxation.
\begin{figure}
%%% plot_sim_mean_field_sweeney.py
\includegraphics[width=\textwidth]{../../plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates (A) and thresholds (B) of excitatory population of 400 Neurons and different values of $\mathrm{\alpha}$ (see equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}).}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}
These findings led us to the conclusion that---apart from the special case $\mathrm{\alpha=1}$, which is the equivalent to instantaneous diffusion---a mean-field model describing the diffusive signal as a mixture of individual activity and all neurons is not suitable for explaining the existence of stable heterogeneous distributions of rates for a broad range of diffusion constants. We did not intent to debase the validity of results of the mean-field model of Sweeney et al. However, we could show that the steady-state solution of this simplified model will result in the same firing rate $\phi_0$ for all neurons, given any value of $\mathrm{\alpha}$ except $\mathrm{\alpha=1}$. This can be seen by setting the left hand side of Equation \eqref{diff_hom_simpl_sweeney} to $0$ (which is necessarily the case in the steady state) and rearranging the equation:
\begin{equation}
(\alpha-1)\frac{\phi_i - \phi_0}{\phi_i} = \alpha \frac{\langle \phi \rangle - \phi_0}{\langle \phi \rangle}
\label{diff_hom_simpl_sweeney_2}
\end{equation}
The right term of the equation is the same for all neurons. Since the left term is monotonically increasing as a function of $\phi_i$ as long as $\mathrm{\alpha<1}$, only one specific solution $\phi_i = \Phi$ for all $i$ exists that equals the given term on the right. This implies $\langle \phi \rangle = \Phi$. Thus,
\begin{equation}
(\alpha-1)\frac{\Phi - \phi_0}{\Phi} = \alpha \frac{\Phi - \phi_0}{\Phi}
\label{diff_hom_simpl_sweeney_3}
\end{equation}
which is only fulfilled for $\Phi = \phi_0$.

@@

Equation \eqref{simple_NO_dyn_with_diff_equil} can be rewritten as
\begin{equation}
\left( \nabla^2 + \left( i\sqrt{\frac{\lambda}{D}}\right)^2\right) NO = \sum_{i} \delta^2(\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i}) \frac{- \gamma \cdot r_i}{D}
\label{simple_NO_dyn_with_diff_equil_helmholtz}
\end{equation}
which is a two-dimensional Helmholtz equation with a superposition of rescaled Dirac functions. The solution of $NO$ is thus composed of a superposition of shifted and scaled versions of the Green's function of the differential operator on the left hand side of the equation \cite{EncyMath}. For each delta function $\delta^2(\mathbf{x}-\mathbf{x}_i)$, the solution is
\begin{equation}
NO_i(\mathbf{x}) = \frac{r_i \gamma}{2 \pi D} K_0 \left(|\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i}|\sqrt{\frac{\lambda}{D}} \right) \equiv r_i \psi_{\rm point}(|\mathbf{x}-\mathbf{x}_{\mathrm{ neur.},i}|)
\label{solution_diff_equil_bessel}
\end{equation}
where $\mathrm{K_0}$ is the zeroth modified Bessel function of the second kind \cite{Helmholtz_Solution_2d}. This solution reveals a fundamental problem of modeling the sources of NO-production as point sources: the fact that $\mathrm{K_0(x)}$ diverges to infinity for $\mathrm{x\rightarrow 0}$. It is merely due to the finite density of the numeric grid used for the simulation of diffusion that allows for a finite target value of concentration. Note that this problem only occurs in the two- or three-dimensional version of the differential equation, whereas in one dimension, the fundamental solution can be expressed as an exponential function with respect to the distance to the origin, resulting in a well-defined finite value at $\mathrm{x=0}$.

@@

By simply calculating all matrix elements of $\mathrm{\hat{\psi}}$ by means of Equation \eqref{interact_matrix_elements}, one would neglect the finite boundaries of the system, which would cause neurons close to the edge to ``bleed" into empty space. This in turn would cause the solution of \eqref{NO_equil_cond_interact_matrix_operator} to contain an over-representation of high firing rates since close-to-the-edge neurons would need to compensate for their lack of neighbors. Thus, we had to account for the boundary conditions used in the network simulation. We simulated the network mostly with Neumann boundary conditions as well as periodic boundaries, as described in section \ref{methods}. Both types can be modeled by extending the neurons' population through spatially shifted and mirrored versions of the base population (see Figure \ref{Bound_Cond_Patches}):
\begin{itemize}
\item Periodic boundary conditions are induced by copies of the neurons' positions shifted by $L\cdot (n_x,n_y)$, $n_x,n_y \in \mathbb{Z}$.
\item Zero flux through the boundaries can be achieved by copied positions being shifted by $L\cdot (n_x,n_y)$, $n_x,n_y \in \mathbb{Z}$ \textit{and} mirrored in $x$($y$)-direction if $n_x$($n_y$) is odd.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{../../plots/Boundary_Cond_Sketch.png}
\end{center}
\caption{Sketch for patches of copied (and mirrored) positions for periodic (A) and Neumann (B) boundary conditions.}
\label{Bound_Cond_Patches}
\end{figure}
Therefore, the entries of the operator for periodic boundary conditions $\hat{\psi}_{\rm per.}$ could be calculated by
\begin{equation}
\psi_{ij,\rm per.} = \sum_{n_x,n_y \in \mathbb{Z}} \psi_{\rm approx.} (|\mathbf{x}_{\mathrm{ neur.},i}-\mathbf{x}_{\mathrm{ neur.},i} - L \cdot (n_x,n_y)|) \; .
\label{Psi_entries_periodic_bound}
\end{equation}
For the Neumann boundary condition, we found
\begin{align}
\psi_{ij,\rm neum.} &= \sum_{n_x,n_y \in \mathbb{Z}} \psi_{\rm approx.} \left(|\mathbf{x}_{\mathrm{ neur.},i}-M(\mathbf{x}_{\mathrm{neur.},i}) - L \cdot (n_x,n_y)| \right) \label{Psi_entries_neumann_bound} \\
M(\mathbf{x}) &\equiv  
 \begin{pmatrix}
  (-1)^{n_x} & 0\\
  0 & (-1)^{n_y}
 \end{pmatrix}
\mathbf{x} + L \cdot (mod(\left| n_x\right|,2),mod(\left| n_y\right|,2)) \; . \label{Mirror_Operator}
\end{align}
Note that the shift was applied to the second position. Theoretically, $n_x$,$n_y$ are to be iterated over all integers and in this case it is irrelevant whether to shift the first or the second position. We obviously had to limit the amount of elements of the sum for our calculations, such that copies that are shifted further away can be neglected. For the sake of performance the number of terms in the sum was adjusted to suffice for the given diffusion constant.

@@

These observations naturally led us to the question of how the correlation between predicted and measured firing rates behaves in between the aforementioned limits. Especially, we were interested in the range of the diffusion constant for which our model provides a good description of the full spiking network's activity. Figure \ref{Corr_Coeff_vs_D} depicts the Pearson correlation coefficient of $\mathrm{f_{matrix \\, predict}}$ and $\mathrm{f_{sim}}$ against the diffusion constant used in the simulation. A relatively high correlation was obtained for a wide range of diffusion constants. However, we could see a general decline of correlation for larger diffusion constants. This trend was in line with the aforementioned limit of instant diffusion, namely a complete decorrelation between prediction and measurement, as well as the good agreement for $\mathrm{D= 10\, \mu m^2 ms^{-1}}$ shown in Figure \ref{Fir_Rate_Dist_Sim_vs_Mat}.
